---
title: "Customer Lifetime Value — Netflix Original Series"
author: "Gianni Hrousis | MKT 326 | Lehigh University"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    code_folding: show
    df_print: kable
    css: custom.css
    includes:
      before_body: _back_link.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
```

---

## 1. What Is CLV and Why Does It Justify Marketing Investments?

Customer Lifetime Value (CLV) is the net present value of all future profits a firm expects to earn from a single customer over the entire duration of the relationship. Rather than treating a marketing spend as a one-time cost, CLV reframes it as an investment with a measurable return: a dollar spent today is justified if and only if the discounted future profit it generates exceeds that dollar. For Netflix, this logic is particularly powerful when applied to content development — a new original series is not simply an entertainment product, it is a retention mechanism. If the series converts non-watchers to watchers, and if watchers retain at meaningfully higher rates than non-watchers, then the CLV difference between the two segments sets a hard ceiling on how much Netflix can rationally invest in that series. In short, CLV gives marketers a defensible, quantitative budget: spend up to the aggregate CLV gain across the affected customer base, and the investment pays for itself.

---

## 2. Parameters


```{r parameters}
# Revenue and cost structure
annual_revenue     <- 11 * 12        # $132 per customer per year
cogs_pct           <- 0.60           # 60% cost of goods sold
marketing_cost     <- 5              # $5 per customer per year
discount_rate      <- 0.10           # 10% annual discount rate
years              <- 9              # projection horizon

# Contribution margin per customer per year
cm <- annual_revenue * (1 - cogs_pct) - marketing_cost
cat("Annual contribution margin per customer: $", round(cm, 2), "\n")

# US customer base
total_us_customers <- 52.77e6
non_watcher_share  <- 0.20
adoption_rate      <- 0.40

# Attrition rates by year (provided in assignment)
attrition_watcher    <- c(0.13, 0.17, 0.19, 0.18, 0.19, 0.17, 0.16, 0.15, 0.14)
attrition_nonwatcher <- c(0.20, 0.26, 0.30, 0.24, 0.29, 0.25, 0.25, 0.21, 0.19)
```

---

## 3. Year-by-Year Retention and Discounted Contributions

Retention in each year is the cumulative probability of a customer still being active — computed as the product of all prior survival rates. Each year's contribution margin is discounted using a mid-year convention (`year - 0.5` as the exponent), which is appropriate for subscription businesses where customers churn throughout the year rather than only at year-end.

```{r clv-table}
# Compute retention (cumulative survival) for each segment
retention_w  <- cumprod(1 - attrition_watcher)
retention_nw <- cumprod(1 - attrition_nonwatcher)

# Discounted contribution per year
disc_cm_w  <- retention_w  * cm / (1 + discount_rate)^((1:years) - 0.5)
disc_cm_nw <- retention_nw * cm / (1 + discount_rate)^((1:years) - 0.5)

# Year-by-year table
clv_table <- tibble(
  Year                    = 1:years,
  `Retention — Watcher`   = round(retention_w,  4),
  `Retention — Non-Watcher` = round(retention_nw, 4),
  `Disc. CM — Watcher ($)`  = round(disc_cm_w,  2),
  `Disc. CM — Non-Watcher ($)` = round(disc_cm_nw, 2),
  `Cum. CLV — Watcher ($)`    = round(cumsum(disc_cm_w),  2),
  `Cum. CLV — Non-Watcher ($)` = round(cumsum(disc_cm_nw), 2)
)

clv_table %>%
  kable(caption = "Year-by-Year Retention, Discounted Contribution, and Cumulative CLV") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  column_spec(6, bold = TRUE, color = "#2c7bb6") %>%
  column_spec(7, bold = TRUE, color = "#d7191c")
```

---

## 4. CLV Summary

```{r clv-summary}
clv_watcher    <- sum(disc_cm_w)
clv_nonwatcher <- sum(disc_cm_nw)
clv_gap        <- clv_watcher - clv_nonwatcher

cat("CLV — Watcher:      $", round(clv_watcher, 2), "\n")
cat("CLV — Non-Watcher:  $", round(clv_nonwatcher, 2), "\n")
cat("CLV Gap:            $", round(clv_gap, 2), " (", round(clv_gap / clv_nonwatcher * 100, 1), "% premium)\n", sep = "")
```

**Interpretation:** A customer who watches Netflix original series is worth **\$`r round(clv_watcher, 2)`** over nine years, compared to **\$`r round(clv_nonwatcher, 2)`** for a non-watcher — a **\$`r round(clv_gap, 2)` (≈`r round(clv_gap/clv_nonwatcher*100,1)`%) premium**. The gap is primarily driven by attrition differences in the first three years: watchers lose 13–19% annually, while non-watchers hemorrhage 20–30% in those same early years. Because CLV compounds over time, early-year retention differences translate into a disproportionately large lifetime value gap.

---

## 5. Maximum Investment in New Original Series

If Netflix develops a new original series that converts non-watchers to watchers at a 40% adoption rate, the maximum rational investment is:

$$\text{Max Investment} = \text{Non-Watchers} \times \text{Adoption Rate} \times \Delta\text{CLV}$$

```{r max-investment}
non_watchers    <- total_us_customers * non_watcher_share
converts        <- non_watchers * adoption_rate
max_investment  <- converts * clv_gap

cat("US non-watchers:         ", scales::comma(non_watchers), "\n")
cat("Converts at 40%:         ", scales::comma(converts), "\n")
cat("Max investment ($M):     $", round(max_investment / 1e6, 0), "M\n", sep = "")
```

**Netflix can justify spending up to \$`r round(max_investment/1e6,0)` million** on a new original series under the 40% adoption assumption. If the series costs less than this threshold, it is CLV-positive — the lifetime value gained from converted customers exceeds the production investment.

---

## 6. Sensitivity Analysis: Adoption Rate

The $363M ceiling depends entirely on the 40% adoption assumption. The table below shows how the maximum rational investment changes across a range of plausible adoption rates.

```{r sensitivity}
adoption_scenarios <- c(0.20, 0.30, 0.40, 0.50, 0.60)

sensitivity <- tibble(
  `Adoption Rate` = scales::percent(adoption_scenarios),
  `Converts (M)`  = round(non_watchers * adoption_scenarios / 1e6, 2),
  `Max Investment ($M)` = round(non_watchers * adoption_scenarios * clv_gap / 1e6, 0)
)

sensitivity %>%
  kable(caption = "Sensitivity of Maximum Investment to Adoption Rate Assumption") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  row_spec(which(adoption_scenarios == 0.40), bold = TRUE, background = "#eaf4fb")
```

**Interpretation:** Even at a conservative 20% adoption rate, Netflix can justify spending **\$`r round(non_watchers * 0.20 * clv_gap / 1e6, 0)`M** on a new series. At the 40% base case it is \$363M, consistent with reported budgets for flagship productions like *Stranger Things* (\~\$300M across multiple seasons) and *The Crown* (\~\$130M per season). This confirms the CLV framework produces economically credible budget ceilings, not theoretical abstractions.

---

## 7. Retention Trajectory Visualization

```{r retention-plot, fig.width=8, fig.height=5}
tibble(
  Year         = rep(0:years, 2),
  Retention    = c(1, retention_w, 1, retention_nw),
  Segment      = rep(c("Watcher", "Non-Watcher"), each = years + 1)
) %>%
  ggplot(aes(x = Year, y = Retention, color = Segment, group = Segment)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.5) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  scale_color_manual(values = c("Watcher" = "#2c7bb6", "Non-Watcher" = "#d7191c")) +
  labs(
    title    = "Netflix Customer Retention Over 9 Years",
    subtitle = "Original series watchers retain at dramatically higher rates",
    x        = "Year",
    y        = "Cumulative Retention (%)",
    color    = "Segment"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "top")
```

---

## 8. Managerial Recommendation

CLV analysis provides Netflix with three actionable conclusions:

1. **Invest in anything that moves non-watchers to watchers.** The \$86 CLV gap means every converted customer is worth nearly \$86 more in discounted lifetime profit. At scale, this is transformational.

2. **Prioritize early retention.** The year-by-year table shows that the CLV advantage is almost entirely established in years 1–3. Strategies that reduce early churn (better onboarding, personalized recommendations, binge-worthy season premieres) compound dramatically over the customer's lifetime.

3. **Use \$363M as a portfolio-level budget ceiling, not a per-show target.** The sensitivity table shows the ceiling is highly sensitive to adoption. Netflix should apply this framework across its content slate to allocate budget to series with the highest estimated conversion potential, not simply the highest production value.


